Better JavaScript Runtime Understanding by Automated Function Name Extraction

### 2011-01-20 trunk/papers/issta2011/sigproc-sp.tex

\usepackage{graphicx}
\usepackage{subfigure}

@@

\maketitle
\begin{abstract}
Understanding JavaScript code due to its dynamic, weakly-typed nature is complicated. Developers usually understand JavaScript programs by running them and examining their elements at runtime. However, understanding concrete values, particularly user-defined and function objects, is not always straightforward. Function names can aid in this regard; they can appear as the constructor name in object representation or as the functions identifier in callstack. Unfortunately, only a very low proportion (less than 10\%) of JavaScript functions are named by developers in the first place. As a solution to this issue, we propose an approach for automated JavaScript function naming based on source code analysis. We applied our approach on several JavaScript projects and the resutls are very promising.

@@

Developers usually understand JavaScript code by examining the running program in debuggers. At runtime, concrete values are available and the developer can directly check and understand object structures and control flows. Understanding a concrete value, particulary for non-primitive values such as objects and functions, is not always straightforward. To facilitate understanding these complex values at the first place, debuggers show a summary of these objects. For example, in case of user-defined objects, the name of object's constructor can be very helpful in the object summary. It can work like a class name and summarizes the object structure.

Functions are central to program comprehension in JavaScript. They are first-class objects that are used for different purposes by developers; 
they may be used as an object  constructor, a clousure scope(module) or even  passed as an argument in a function call. However, there is a subtle issue with JavaScript functions that makes working with them quite difficult: functions can be defined and created with no name or identifier. This problem particularly shows up in debuggers \cite{Zaytsev}. The function name is necessary for referring and recalling the function's functionality, but only a small proportion (less than 10\%) of JavaScript functions are named.

@@

@@

{\small\texttt{Function}} objects have no name by default; as the {\small\texttt{Function}} constructor doesn't receive any name. According to ECMA specification, a function name (identifier) can be specified after the {\small\texttt{function}} operator (It is mandatory for function declerations but optional for function expressions) \cite{ECMA}. However, the usage of function identifiers is limited. The reason is that, function names, in addition to naming the function, has an effect on available variables in the scopes. In the case of function declearations, the function object is assigned to a variable in the function scope, but in the case of function expressions, the function object is available only inside the function. The confusing behavior of function identifiers get worse by function statements introduced by some browser such as Firfox. Function statements are like function expressions, but the function object is available in the entire function scope like function declarations \cite{Zaytsev}. Therefore, developers often prefer to omit the function identifier in the function statement to avoid its confusing impact.

@@

Figure~\ref{js-code} demonstrates an excerpt of a JavaScript code. We set a breakpoint on line 13 and examin the runtime elements at this breakpoint by two JavaScript debuggers, Google Chrome and Firebug (Figure~\ref{debuggers-objects}). Two objects assigned to variables {\small\texttt{foo}} and {\small\texttt{bar}} are constructed by two different constructors: {\small\texttt{FOO}} and {\small\texttt{BAR}}. However, Google Chrome debugger shows the general class of {\small\texttt{Object}} as the summary for both objects. The developer has to expand the object nodes to recognize their similarities and differences. Firebug classify objects in the same general class, but includes some of the object properties in the summary. These additional properties may hint develooper about the object structures. {\small\texttt{FOO}} and {\small\texttt{BAR}} definitions at lines 19 and 22 explain this behaviour; the function statements has no explicit name (identifier), therefore debuggers considered them as \textit{anonymous} functions.

@@

\subsection{Callstack View}
To illustrate the second issue, we pause the program (Figure~\ref{js-code}) at line 15 by a breakpoint. Figure~\ref{debuggers-callstack} shows how the program callstack is displayed in Google Chrome debuggers and Firebug. The differences in the number of frames and line numbers between two callstacks are due to dissimilar event handling implementations in the underlying platforms. Google chrome shows {\small\texttt{anonymouse}} as the name of all functions excepting the first one. Firebug performs better by guessing three function names but it still fails in one case. In these cases, the information provided by debugger is useless and the developer has to locate the function source to understand or recall the function behaviour.

@@

\section{Automated Funcation Naming}
    To tackle the aforementioned function naming issues, we propose automated function naming by anlysing the source code. A function name must help the developer to
distinguish the function from other functions with different behaviours. The behaviour of a function is dependent to its source code, enclosing scopes (because of global and closure variables), the object bound to {\small\texttt{this}}-depending how it is called- and passed arguments. The first two parameters can be almost known from the source code, while the two last parameters are usually not known until the function call at runtime. Therefore, to name a function object, it is enough to name its \textit{function body}. 
    A function body can be uniquly identified by its location in the source file. A naive proposal for the function name for a function body is a combination of the function file name and line number. Although it may work as an identifier, but it does not assist the develper to recall or understand the function behaviour.

In contrast, we use names in the source code to name nameless functions. 
    Once a function object is created, one of these cases would happen:
    \begin{enumerate}
		\item the function object is assigned to a variable or one of its properties.
		\item it is assigned to a property of a new object.
		\item it is assigned to a new array index.
		\item it is called.
		\item it is returned from a function call.
		\item it is passed as an argument to a function.
		\end{enumerate}

To choose the most appropriate name for a function which has the characteristics defined in the previous section, we consider two cases. First, assigning a global name to the function which identifies the function in the whole program. Second, naming a function in a call stack.

A function object in JavaScript can be defined by \texttt{function \textit{identifier} (args)\{...\}} or \texttt{new Function(args, source)}.
\subsection{A Global Name for a Function Object}

%\begin{center}
    %\begin{tabular}{ | l | l | l | p{5cm} |}
    %\hline
    %& Type & Sample & Description \\ \hline
    %a & Named & function foo(){...} & The function is named. \\ \hline
    %a & Named & function foo(){...} & The function is named. \\ \hline
    %a & Named & function foo(){...} & The function is named. \\ 
    %\hline
    %\end{tabular}
%\end{center}

\begin{figure}[htp]
\begin{verbatim}
(a) Case 1:
 function foo(){ ... }

(b) Case 2:
 var foo = function { ... }

(c) Case 3:
 var foo = {
     bar : function { ...}
 }

(d) Case 4:

 var foo = function(){
           ....
           return function(){ ... }
 }()

(e) Case 5:
 var foo = function(){ 
 					 return ...;
 }()

(f) Case 6:
 var foo = function(){

 }()();

(g)
 foo(function(){ ... });

(h)
 array = [...., function(){},...]

(i)
 array[i] = function(){};

(j)
 eval(); & new Function("");

\end{verbatim}
\caption{Different cases of function object creation.}
\label{fig:functionCreation}
\end{figure}

\subsection{Naming Functions in Call Stack}
%new Function ([arg1[, arg2[, ... argN]],] functionBody)
%this[ optall.queue === false ? "each" : "queue" ](function() {})

>>>>

\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{array}

@@

\maketitle
\begin{abstract}
Understanding JavaScript code due to its dynamic, weakly-typed nature is complicated. Developers usually understand JavaScript programs by running them and examining their elements at runtime. However, understanding concrete values, particularly user-defined and function objects, is not always straightforward. Function names can aid in this regard; they can appear as the constructor name in object representation or as the functions identifier in callstack. Unfortunately, only a very low proportion (less than 5\%) of JavaScript functions are named by developers in the first place. As a solution to this issue, we propose an approach for automated JavaScript function naming based on source code analysis. We applied our approach on several JavaScript projects and the resutls are very promising.

@@

Developers usually understand JavaScript code by examining the running program in debuggers. At runtime, concrete values are available and the developer can directly check and understand object structures and control flows. Understanding a concrete value, particulary for non-primitive values such as objects and functions, is not always straightforward. To facilitate understanding these complex values in the first place, debuggers show a summary of these objects. For example, in case of user-defined objects, the name of object's constructor can be very helpful in the object summary. It can work like a class name and summarizes the object structure.

Functions are central to program comprehension in JavaScript. They are first-class objects that are used for different purposes by developers; 
they may be used as an object  constructor, a clousure scope(module) or even  passed as an argument in a function call. However, there is a subtle issue with JavaScript functions that makes working with them quite difficult: functions can be defined and created with no name or identifier. This problem particularly shows up in debuggers \cite{Zaytsev}. The function name is necessary for referring and recalling the function's functionality, but only a small proportion (less than 5\%) of JavaScript functions are named.

@@

\begin{table}
\centering
\caption{The total number of functions and the number of named functions in five famous JavaScript projects.}
  \begin{tabular}{ | l | l | c | c | c |}
  \hline
   Project & Description & Total & Named \\ 
  \hline 
   jQuery & JavaScript Library & 422 & 23 \\ 
  \hline 
   Prototype & JavaScript Framework & 645 & 203 \\ 
  \hline 
   Closure & Google Web Library & 9195 & 208 \\ 
  \hline 
   YUI &  Yahoo UI Library & 22346 & 922 \\ 
  \hline 
   Firebug & Web Development Tool & 3424 & 406 \\ 
  \hline 
  \end{tabular}
 \label{functions} 
\end{table}

@@

{\small\texttt{Function}} objects have no name by default; as the {\small\texttt{Function}} constructor doesn't receive any name. According to ECMA specification, a function name (identifier) can be specified after the {\small\texttt{function}} operator (It is mandatory for function declerations but optional for function expressions) \cite{ECMA}. However, the usage of function identifiers is limited. The reason is that, function names, in addition to naming the function, has an effect on available variables in the scopes. In the case of function declearations, the function object is assigned to a variable in the function scope, but in the case of function expressions, the function object is available only inside the function. The confusing behavior of function identifiers get worse by function statements introduced by some browsers such as Firfox. Function statements are like function expressions, but the function object is available in the entire function scope like function declarations \cite{Zaytsev}. Therefore, developers often prefer to omit the function identifier in the function statement to avoid its confusing impact.

To support the above argument, we analyzed the source code of five famous JavaScript projects. For every project, the total number of functions and the number of functions with name are collected (Table~\ref{functions}). The ratio of the number of functions with names to the total number of functions is less than 13 percent in all projects excepting Prototype and the average ratio is less than 5 percent. Our analysis does not include the functions defined by dynamically created scripts evaluated by {\small\texttt{eval}} function.

@@

Figure~\ref{js-code} demonstrates an excerpt of a JavaScript code. We set a breakpoint on line 13 and examin the runtime elements at this breakpoint by two JavaScript debuggers, Google Chrome and Firebug (Figure~\ref{debuggers-objects}). Two objects assigned to variables {\small\texttt{foo}} and {\small\texttt{bar}} are constructed by two different constructors: {\small\texttt{FOO}} and {\small\texttt{BAR}}. However, Google Chrome debugger shows the general class of {\small\texttt{Object}} as the summary for both objects. The developer has to expand the object nodes to recognize their similarities and differences. Firebug classify objects in the same general class, but includes some of the object properties in the summary. These additional properties may hint develooper about the object structures. {\small\texttt{FOO}} and {\small\texttt{BAR}} definitions at lines 19 and 22 explain the debuggers' behaviour; the function statements has no explicit name (identifier), therefore debuggers considered them as \textit{anonymous} functions.

@@

\subsection{Callstack View}
To illustrate the second issue, we pause the program (Figure~\ref{js-code}) at line 15 by a breakpoint. Figure~\ref{debuggers-callstack} shows how the program callstack is displayed in Google Chrome debugger and Firebug. The differences in the number of frames and line numbers between two callstacks are due to dissimilar event handling implementations in the underlying platforms. Google chrome shows {\small\texttt{anonymouse}} as the name of all functions excepting the first one. Firebug performs better by guessing three function names but it still fails in one case. In these cases, the information provided by debugger is useless and the developer has to locate the function source to understand or recall the function behaviour.

@@

\section{Automated Funcation Naming}
    To tackle the aforementioned function naming issues, we propose automated function naming by anlysing the source code. A function name must help the developer to
distinguish the function from other functions with different behaviours. The behaviour of a function is dependent to its source code, enclosing scopes (e.g., global and closure variables), the object bound to {\small\texttt{this}}-depending on the way the function is called-and passed arguments. The first two parameters can be almost known from the source code, while the two last parameters are usually not known until the function call at runtime. Therefore, to name a function object, it is enough to name its \textit{function body}. 
    A function body can be uniquly identified by its location in the source file. A naive proposal for the function name for a function body is a combination of the function file name and line number. Although it may work as an identifier, but it does not assist the develper to recall or understand the function behaviour. Morover, A proposed function name should be close to the name a developer would choose for the function.

Once a function object is created, one of these cases would happen: the function object is
    \begin{enumerate}
		\item assigned to a property of a new object in an object literal.
		\item assigned to a new array index in an array literal.
		\item assigned to an object property through direct access.
		\item assigned to an object property through hashmap access by a string.
		\item assigned to an object property through hashmap access by a variable name.
		\item assigned to an object property through hashmap access by a JavaScript expression.
		\item assigned to an array index.
		\item assigned to a variable.
		\item directly called.
		\item returned from a function call.
		\item passed as an argument to a function.
		\end{enumerate}

\begin{table}
\centering
\caption{Different cases of function object creation and usage in JavaScript.}
\scalebox{0.8}{
\renewcommand\arraystretch{2.0}
\begin{tabular}{ | c | l | l | l | m{2.5cm} | l|}
  \hline
   & \multicolumn{4} {| c |}{Description} & \multicolumn{1} {| c |}{Code} \\ 
  \hline

   1 &
   &  
   & \multicolumn{2}{|m{2.8cm}|}{
     \raggedright property of a new object in an object literal.}
   & \{ ..., foo: function()\{...\}, ...\}\\ %{\small\begin{verbatim}{ ..., foo: function(){...}, ...}\end{verbatim}} \\ %
   \cline{0-0}\cline{4-6}

	 2 &
	 & \multirow{8}{*}{\hspace{.2cm}\begin{rotate}{90}assigned to a(n)\end{rotate}}
	 & \multicolumn{2}{|m{2.8cm}|}{
	   new array index in an array literal.} 
   & foo = [..., function()\{...\}, ...] \\ 
   \cline{0-0}\cline{4-6}

   3 &
   & 
   &  
   & direct access.
   & bar.foo = function()\{...\}\\
   \cline{0-0}\cline{5-6}

   4 & \multirow{11}{*}{\hspace{.2cm}\begin{rotate}{90}The function object is\end{rotate}} 
   & 
   & 
   & \raggedright hashmap access by a string. 
   & bar["foo"] = function()\{...\} \\
   \cline{0-0}\cline{5-6}

   5 &
   & 
   & \multirow{4}{*}{\hspace{.2cm}\begin{rotate}{90}object property through\end{rotate}}
   & \raggedright hashmap access by a variable name.
   & bar[foo] = function()\{...\} \\
   \cline{0-0}\cline{5-6}

   6 &
   & 
   & 
   & \raggedright hashmap access by a JavaScript expression.
   & bar["f"+"oo"] = function()\{...\} \\
   \cline{0-0}\cline{4-6}

   7 &
   &
   & \multicolumn{2}{|l|}{
      array index.}
   & foo[0] = function()\{...\} \\
   \cline{0-0}\cline{4-6}

   8 &
   & 
   & \multicolumn{2}{|l|}{
      variable.}
   & foo = function()\{...\} \\
   \cline{0-0}\cline{3-6}

   9 &
   & \multicolumn{3}{|m{3.8cm}|}{
     \raggedright directly called.}
   & function()\{...\}() \\
   \cline{0-0}\cline{3-6}

   10 &
   & \multicolumn{3}{|m{3.5cm}|}{
     \raggedright returned from a function call.}
   & \{... return function()\{...\}\} \\
   \cline{0-0}\cline{3-6}

   11 &
   & \multicolumn{3}{|m{3.5cm}|}{
     \raggedright passed as an argument to a function.}
   & foo(..., function()\{\}, ...) \\
   \hline

  \end{tabular}
  }
\end{table}

%to remove
To illustrate the second issue, we pause the program (Figure~\ref{js-code}) at line 15 by a breakpoint. Figure~\ref{debuggers-callstack} shows how the program callstack is displayed in Google Chrome debuggers and Firebug. The differences in the number of frames and line numbers between two callstacks are due to dissimilar event handling implementations in the underlying platforms. Google chrome shows {\small\texttt{anonymouse}} as the name of all functions excepting the first one. Firebug performs better by guessing three function names but it still fails in one case. In these cases, the information provided by debugger is useless and the developer has to locate the function source to understand or recall the function behaviour.

To illustrate the second issue, we pause the program (Figure~\ref{js-code}) at line 15 by a breakpoint. Figure~\ref{debuggers-callstack} shows how the program callstack is displayed in Google Chrome debuggers and Firebug. The differences in the number of frames and line numbers between two callstacks are due to dissimilar event handling implementations in the underlying platforms. Google chrome shows {\small\texttt{anonymouse}} as the name of all functions excepting the first one. Firebug performs better by guessing three function names but it still fails in one case. In these cases, the information provided by debugger is useless and the developer has to locate the function source to understand or recall the function behaviour.

To illustrate the second issue, we pause the program (Figure~\ref{js-code}) at line 15 by a breakpoint. Figure~\ref{debuggers-callstack} shows how the program callstack is displayed in Google Chrome debuggers and Firebug. The differences in the number of frames and line numbers between two callstacks are due to dissimilar event handling implementations in the underlying platforms. Google chrome shows {\small\texttt{anonymouse}} as the name of all functions excepting the first one. Firebug performs better by guessing three function names but it still fails in one case. In these cases, the information provided by debugger is useless and the developer has to locate the function source to understand or recall the function behaviour.

\subsection{Naming Functions in Call Stack}

%To choose the most appropriate name for a function which has the characteristics defined in the previous section, we consider two cases. First, assigning a global name to the function which identifies the function in the whole program. Second, naming a function in a call stack.

%A function object in JavaScript can be defined by \texttt{function \textit{identifier} (args)\{...\}} or \texttt{new Function(args, source)}.
%\subsection{A Global Name for a Function Object}

%new Function ([arg1[, arg2[, ... argN]],] functionBody)
%this[ optall.queue === false ? "each" : "queue" ](function() {})
